Physcraper: a python package for continual update of evolutionary estimates using the Open Tree of Life

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
intro

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
Phylogenies capture the shared history of organisms and provide key evolutionary
context for our biological observations \cite{dobzhansky1973nothing}.
The Open Tree of Life project (OpenTree)
constructs a comprehensive tree of life by synthesizing published phylogenies
and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
placed in the tree based on taxonomy. Meaning that there is still a lot of work
to do to complete a full phylogenetic tree of all life.

@@

Taxonomic idiosyncrasies across databases represent a huge challenge for automatic
integration of data into phylogenies, which can be addressed with a unified
taxonomy for name standardization.

To achieve this, OpenTree unifies taxonomic
data from various databases \cite{rees2017automated}, including the USA National
Center for Biodiversity Information (NCBI) molecular database GenBank
\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
for connecting data from any biological database that has been integrated to it.

>>>>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
Phylogenies capture the shared history of organisms and provide key evolutionary
context for our biological observations \cite{dobzhansky1973nothing}.
Public biological databases constitute an amazing resource for evolutionary
studies. Connecting existing phylogenies to molecular data that has never been
incorporated into any phylogenetic estimate, geographical location, fossils, and
other data in a reproducible and continuous manner is possible by establishing a
data interoperability framework for biological databases. Here, we introduce
Physcraper, a tool that automates biological database connections
with the main goal of building upon homology
hypotheses that taxon specialists have assessed and deemed appropriate for a
specific phylogenetic scope to update a starting tree and single locus alignments
with public DNA data, with the added capability to connect the lineages represented
in the phylogeny to biological data in public databases.

One of the main challenges for automatic
integration of biological data into phylogenies are varying taxonomic idiosyncrasies
across databases. To address this challenge, the Open Tree of Life project (OpenTree)
unifies taxonomic
data from various databases \cite{rees2017automated}, including the USA National
Center for Biodiversity Information (NCBI) molecular database GenBank
\cite{benson2000genbank}, among others. The OpenTree taxonomy represents a key resource
that remains to be leveraged for connecting data from any biological database that
has been integrated to it.

The Open Tree of Life project (OpenTree)
also constructs a comprehensive tree of life by synthesizing published phylogenies
and taxonomy. OpenTree's "synthetic" tree comprises 2.3 million tips, of which
around 90,000 (87,740) are supported by phylogenies - the remaining 1.4 million taxa are
placed in the tree based on taxonomy. Meaning that there is still a lot of work
to do to complete a full phylogenetic tree of all life.

@@

### 2021-01-07 temp.Rmd
intro

Public biological databases constitute an amazing resource for evolutionary
studies. Connecting existing phylogenies to molecular data that has never been
incorporated into any phylogenetic estimate, geographical location, fossils, and
other data in a reproducible and continuous manner is possible by establishing a
data interoperability framework for biological databases. Here, we introduce
Physcraper, a tool that automates biological database connections
with the main goal of building upon homology
hypotheses that taxon specialists have assessed and deemed appropriate for a
specific phylogenetic scope to update a starting tree and single locus alignments
with public DNA data.

>>>>

\cite{pyron2011phylogeny} PHLAWD, Colubroidea, 334 citations

Large genomic dataset limit the complexity of evolutionary models that can be applied.
Single locus allows for more complex evolutionary models, and therefore might
provide better evolutionary estimates \cite{}.

Single locus super-matrix have provided very good phylogenegic inferences, sometimes even better than genomic based estimates. SL also, allow for construction of large phylogenies and

Line 78- Add extra sentence about advantages of Physcraper's implementation
Physcraper has speciifc advantages over existing pipelines.
